% Report #3: Industry
% Andrey Ziyatdinov
% 15/08/2013

<link rel='stylesheet' href=inc/nv.d3.css>
<link rel='stylesheet' href=inc/rNVD3.css>
<script type='text/javascript' src=inc/jquery-1.8.2.min.js></script>
<script type='text/javascript' src=inc/d3.v3.min.js></script>
<script type='text/javascript' src=inc/nv.d3.min-new.js></script>
<script type='text/javascript' src=inc/fisheye.js></script>


```{r setup, include = FALSE}
# set global chunk options
opts_chunk$set(fig.path = 'figure/', cache.path = 'cache/', echo = FALSE, cache = TRUE, 
  fig.width = 9, fig.height = 4.5, dev = 'svg',
  warning = FALSE, message = FALSE)
# upload images automatically? (identity/imgur_upload)
opts_knit$set(upload.fun = identity)
```

```{r include_hidden, echo = FALSE, results = "hide", warning = FALSE, message = FALSE}
library(xtable)
library(ascii)
options(asciiType = 'pandoc')

library(plyr)
library(reshape)

library(maps)

library(hexbin)
library(ggplot2)
library(wq)
library(grid)
library(RColorBrewer)

library(rCharts)

library(devtools)
load_all("~/git/variani/NLPPeopleStat/pck/")
```

```{r params}
include_assets <- TRUE
cdn <- FALSE
```

```{r data}
load("../../R/rdata/postings.RData")

# Subcat
postings <- within(postings, {
  Sub.category[Sub.category == "Consultants/experts"] <- "Consultants/Experts"
  Sub.category <- droplevels(Sub.category)
})

# level names
levels(postings$Sub.category)[1:7] <- c("Researchers", "Developers", "Consultants", 
  "Managers", "Others", "Comp. Linguists", "Internships")
# level order  
lev <- levels(postings$Sub.category)
postings$Sub.category <- factor(postings$Sub.category, levels = lev[c(1:2, 6, 3:4, 5, 8:9)])

# Exp
uval <- sort(unique(as.character(postings$Level.of.experience.years)))
#[1] "Entry Level (Less than 2 years of Experience)"
#[2] "Executive (SVP, EVP, VP etc)"                 
#[3] "Internship"                                   
#[4] "Management (Manager/Director)"                
#[5] "Mid Career (2+ years of experience)"          
#[6] "Not Specified"                                
#[7] "Senior (5+ years of experience)" 
postings <- mutate(postings,
  Level.of.experience.years = factor(Level.of.experience.years,
    levels = uval[c(1, 5, 7, 4, 2, 6)]))
    
postings$Rec <- "Direct"
ind <- grep("Rec\\.|rec\\.", postings$Company)
postings$Rec[ind] <- "Rec."
postings$Rec <- factor(postings$Rec)
    
postings2 <- subset(postings, Company.type %in% "Industry")
```

# Introduction

The Report No 2 was dedicated to the study of the educational level and its influence to the career opportunities.
The educational level was mainly considered in relation to the research positions.
We expect that the job positions, which are not-research and from the industry sector, 
have many other factors rather than the educational level.

In this report we will study descriptors of the job postings (including the educational level)
that play the important role in the industry sector.

The work flow of the analysis will be the following.

* Take the postings data that belong to the industry sector.
* Select a set of subcategories, which have enough data, and do the analysis only for these subgroups.
* Check the following list of factors: educational level, experience, type of contract, recruiting or not, technical skills.
* Before doing the statistics, filter the dataset for removing any small group of the postings.

## Derived variables

### Rec

We derived a new variable `Rec` based on the value of `Company` variable
(the pattern `Rec.` is given or not).
This variable encodes two levels, `Rec` and `Direct`, 
and means whether a posting comes from a recruiting company or not.

```{r, results = "asis"}
tab <- ddply(postings2, "Rec", summarize, Freq = length(N))
ascii(tab, format = c("s", "d"), na.print = "N/a")
```

### Technical Skill

We derived a new variable for description of the technical skills
mentioned in the text of `Qualifications` variable.
In particular, we extracted patterns like `Cpp` or `Java`.

The definition of such kind of variables is under development,
and here we did it the first time and by simply using `grep` command.

# Time Plots

The data used in the report were collected from `r as.character(min(postings$Post.date, na.rm = TRUE))`
till `r as.character(max(postings$Post.date, na.rm = TRUE))`.

## Number of postings by Subcategory ~ time 

The following figure shows the dynamics of the number of postings over time grouped by subcategory.

Note that the presented number describe the statistics only in the industry sector.

```{r results = 'asis', comment = NA}
# prepare data
lev <- levels(postings$Sub.category)
lev <- lev[!is.na(lev)] # remove NA
  
sf <- df_to_sf_month(postings, "Sub.category", lev)

sf <- mutate(sf,
  month.year = as.character(month.year),
  Date = as.Date(paste("1", sf$month.year, sep = "-"), format = "%d-%b-%y")
)
# plot: Category ~ time
p1 <- nPlot(value ~ Date, group =  'variable', data = sf, type = 'stackedAreaChart', id = 'chart')

p1$xAxis( tickFormat="#!function(d) {return d3.time.format('%b %Y')(new Date( d * 86400000 ));}!#" )
p1$chart(color = brewer.pal(10, "Paired"))

p1$print('p1', include_assets = include_assets , cdn = cdn)
```

```{r}
subcat.levels <- levels(postings$Sub.category)[1:5]
```

If one switch to the *Stream* view of the stacked area plot,
it will be possible to select the most representative subcategories visually.
These subcategories include: `r paste(subcat.levels, collapse = ", ")`.

From now, we will consider only these 6 subcategories.

# Filters

We apply a set of filters for removing the records which are not representative.

```{r}
edu.levels <- levels(postings$Educational.level)[1:4]
exp.levels <- levels(postings$Level.of.experience.years)[1:5]
contract.levels <- levels(postings$Type.of.contract)[c(1:2, 4)]
```

## Educational level

```{r, results = "asis"}
tab <- ddply(postings2, "Educational.level", summarize, Freq = length(N))
ascii(tab, format = c("s", "d"), na.print = "N/a")
```

Educational levels which pass the filter: `r paste(edu.levels, collapse = ", ")`.

## Experience

```{r, results = "asis"}
tab <- ddply(postings2, "Level.of.experience.years", summarize, Freq = length(N))
ascii(tab, format = c("s", "d"), na.print = "N/a")
```

Experience levels which pass the filter: `r paste(exp.levels, collapse = ", ")`.

## Type of Contract

```{r, results = "asis"}
tab <- ddply(postings2, "Type.of.contract", summarize, Freq = length(N))
ascii(tab, format = c("s", "d"), na.print = "N/a")
```

Contract levels which pass the filter: `r paste(contract.levels, collapse = ", ")`.

## Application of Filters

```{r}
postings3 <- subset(postings2, Educational.level %in% edu.levels &
  Level.of.experience.years %in% exp.levels &
  Sub.category %in% subcat.levels & 
  Type.of.contract %in% contract.levels)
postings3 <- droplevels(postings3)
```

Further in the next section of the analysis,
we applied the filters described above separately.
The reason is that the number of records is reduced dramatically from `r nrow(postings2)` to `r nrow(postings3)`
after we apply all the three filters together.
That can be explained with a lot of missing data for each variable per filter.

# Educational level ~ Subcategory

```{r}
df <- subset(postings2, Sub.category %in% subcat.levels &
  Educational.level %in% edu.levels)
df <- droplevels(df)  
```

Before computing the statistics, we applied the filter for the educational level,
that leads to a new dataset with the number of records equal to `r nrow(df)`.

The following figure shows the distribution of the postings across the educational level
for each subcategory.

```{r, results = "asis"}
sf <- ddply(df, c("Sub.category", "Educational.level"), summarize, Freq = length(N))

p2 <- nPlot(Freq ~ Sub.category, group = 'Educational.level', data = sf, type = 'multiBarChart')
p2$print('p2', include_assets = include_assets , cdn = cdn)
```

Observations.

* The PhD is mostly required for Researchers.
* For all the subcategories except Researches and Computational Linguists,  
  the major number of job offers are proposed for Bachelors.

# Experience ~ Subcategory

```{r}
df <- subset(postings2, Sub.category %in% subcat.levels &
  Level.of.experience.years %in% exp.levels)
df <- droplevels(df)  

df <- within(df, {
  Level.of.experience.years[Level.of.experience.years %in% "Executive (SVP, EVP, VP etc)"] <- "Management (Manager/Director)"
  Level.of.experience.years <- droplevels(Level.of.experience.years)
})
```

Before computing the statistics, we applied the filter for the experience level,
that leads to a new dataset with the number of records equal to `r nrow(df)`.


The figure given belows shows the distribution of the postings across the experience levels
for each subcategory.

```{r, results = "asis"}
sf <- ddply(df, c("Sub.category", "Level.of.experience.years"), summarize, Freq = length(N), .drop = FALSE)

p3 <- nPlot(Freq ~ Sub.category, group = 'Level.of.experience.years', data = sf, type = 'multiBarChart')
p3$print('p3', include_assets = include_assets , cdn = cdn)
```

Observations.

* For two the most popular subcategories Researchers and Developers, 
  the majority of the job offers are proposed to the candidates with the mid-experience (+2 years).
* For the category Managers, the specific requirements as Management and Executive are applied.
* The job offers for Computational Linguists are mostly offered to the mid-experiences candidates (+2 years).

# Type of contract ~ Subcategory

The figure given below shows that the stable contracts (FULLTIME and CONTRACT) 
are offered to the most of the jobs for all the subcategories.

```{r}
df <- subset(postings2, Sub.category %in% subcat.levels &
  Type.of.contract %in% contract.levels)
df <- droplevels(df)  
```

Before computing the statistics, we applied the filter for the contract level,
that leads to a new dataset with the number of records equal to `r nrow(df)`.


```{r, results = "asis"}
sf <- ddply(df, c("Sub.category", "Type.of.contract"), summarize, Freq = length(N), .drop = FALSE)

p5 <- nPlot(Freq ~ Sub.category, group = 'Type.of.contract', data = sf, type = 'multiBarChart')
p5$print('p5', include_assets = include_assets , cdn = cdn)
```

# Rec. ~ Subcategory

```{r}
df <- subset(postings2, Sub.category %in% subcat.levels)
df <- droplevels(df)  
```

The following figure presents the distribution of the postings across 
the two types (Rec. and Direct) for each subcategory.

```{r, results = "asis"}
myprop <- function(x) 100 * sum(x) / length(x)

sf <- ddply(df, "Sub.category", summarize, 
  Direct = myprop(Rec %in% "Direct"), 
  Rec = myprop(Rec %in% "Rec.")) 

mf <- melt(sf, id.vars = "Sub.category")

p6 <- nPlot(value ~ Sub.category, group = 'variable', data = mf, type = 'multiBarChart')
p6$print('p6', include_assets = include_assets , cdn = cdn)
```

Observations.

* In the view *Stacked* one can clearly see that 
  the subcategory Managers differs from the others:
  the relatively greater portion of the jobs is offered by the recruiting companies.
  
# Technical skills ~ Subcategory

The figure shows the distribution of the technical skills for each subcategory.

The skill labeled as *Other* means that none of the defined skills (Java, Cpp, Python, Perl, SQL)
was found the this group of the postings. Perhaps, the postings labeled as *Other*
require other skills like R or Matlab.

```{r, results = "asis"}
#> subcat.levels
#[1] "Researchers"     "Developers"      "Comp. Linguists" "Consultants"    
#[5] "Managers"        "Internships"     "Others" 
df <- subset(postings2, Sub.category %in% subcat.levels[c(1:3)])
df <- droplevels(df)  

grep_to_log <- function(pat, x)
{
  1:length(x) %in% grep(pat, x)
}

df <- mutate(df,
  Java = grep_to_log("Java|java", Qualifications),
  Cpp = grep_to_log("C\\+\\+|c\\+\\+", Qualifications),
  Python = grep_to_log("Python|python", Qualifications),
  Perl = grep_to_log("Perl|perl", Qualifications),
  SQL = grep_to_log("sql|Sql|SQL", Qualifications),
  SkillNone = !(Java | Cpp | Python | Perl | SQL)
)  

sf <- ddply(df, "Sub.category", summarize, 
  Java = myprop(Java), Cpp = myprop(Cpp), Python = myprop(Python), 
  Perl = myprop(Perl), SQL = myprop(SQL),
  Other = myprop(SkillNone),
  .drop = FALSE)
mf <- melt(sf, id.vars = "Sub.category")

p4 <- nPlot(value ~ Sub.category, group = 'variable', data = mf, type = 'multiBarChart')
p4$print('p4', include_assets = include_assets , cdn = cdn)
```

Observations.

* For the two most popular subcategoires Researchers and Developers,
  the skill of Java seems to be the most wanted.
* For subcategory Computational Linguists, the skill of Perl is the most popular.
* For subcategories Consultants, Managers and Other, the skill of SQL is the most popular.

# TODO

* The figures for statistics on Contract and Rec are quite trivial.
  It will be better to remove the figure and substitute it with the table or some text.
* It will be important to support the different relation *Job and Requirements* by examples of the job offers.
    * A job posting for Computational Linguists with skill of Perl.
    * A job posting for Managers with special requirements in the experience level.
    
    


